Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-03-09 paper-amia-2018/manuscript.tex
submitted version of amia 2018

We experimented both with traditional classifiers, including Naive Bayes (NB)\cite{pedregosa2011scikit}, Support Vector Machine (SVM)\cite{chang2011libsvm}, K-Nearest Neighbor (KNN)\cite{pedregosa2011scikit} and two variants of Recurrent Neural Networks (RNNs)\cite{bengio1993problem}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{cho2014properties}.

@@

\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns. When we predict the ``new segment'' point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior. Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. GRU\cite{cho2014properties} is a variant of LSTM mathematically represented by the following formula:

\begin{equation}
z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
\label{eq:firstgru}
\end{equation}
\begin{equation}
r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
\label{eq:resetgru}
\end{equation}
\begin{equation}
\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
\label{eq:candidategru}
\end{equation}
\begin{equation}
h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
\label{eq:lastgru}
\end{equation}  
In Eq.~\ref{eq:firstgru}-\ref{eq:lastgru}, $\sigma$ corresponds to sigmoid function and $\odot$ designates an element-wise product. The update gate $z_t$ and reset gate $r_t$ at time step $t$ are computed by the Eq.~(\ref{eq:firstgru}) and~(\ref{eq:resetgru}), where $W_z$, $W_r$, $W_h$, $U_z$, $U_r$, $U_h$ are the weight matrices and $b_z$, $b_h$ and $b_r$ are bias vectors. The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}). Our GRU model includes one hidden layer, output layer, and input layer. We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector. Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used. We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1, and Adam is used for optimization.

>>>>

We experimented both with traditional classifiers, including Naive Bayes (NB)\cite{mccallum1998comparison}, Support Vector Machine (SVM)\cite{cortes1995support}, K-Nearest Neighbor (KNN)\cite{keller1985fuzzy} and two variants of Recurrent Neural Networks (RNNs)\cite{lipton2015critical}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{chung2014empirical}.

@@

\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns. When we predict the ``new segment'' point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior. Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:

\begin{equation}
z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
\label{eq:firstgru}
\end{equation}
\begin{equation}
r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
\label{eq:resetgru}
\end{equation}
\begin{equation}
\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
\label{eq:candidategru}
\end{equation}
\begin{equation}
h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
\label{eq:lastgru}
\end{equation}  
In Eq.~\ref{eq:firstgru}-\ref{eq:lastgru}, $\sigma$ corresponds to sigmoid function and $\odot$ designates an element-wise product. The update gate $z_t$ and reset gate $r_t$ at time step $t$ are computed by the Eq.~(\ref{eq:firstgru}) and~(\ref{eq:resetgru}), where $W_z$, $W_r$, $W_h$, $U_z$, $U_r$, $U_h$ are the weight matrices and $b_z$, $b_h$ and $b_r$ are bias vectors. The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}). Our GRU model includes one hidden layer, output layer, and input layer. We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector. Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used. We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1, and Adam\cite{kingma2014adam} is used for optimization.
